---
title: "S-R Compatibility Task - Model Comparison"
author: 
    name: Milan Ewert, Tim Bax, Niklas Bub, Özge Özenoglu
    email: mewert@uos.de, tbax@uos.de, nbub@uos.de, ooezenoglu@uos.de
date: "July 12, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    highlight: tango
---

In this file, we will compare five models against each other by their posterior predictive densities using the Leave-One-Out Cross-Validation (LOO-CV) method. We justify the model used for analysis based on the outcome of the comparisons below. 

# Preliminaries

## Load libraries

```{r echo = TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
library(brms)

options(mc.cores = parallel::detectCores())
```

## Load raw data 

```{r}

raw_data <- read.csv('https://raw.githubusercontent.com/ooezenoglu/XPlab-2020-SR-Compatibility/master/data/01_pilot/01-raw-data-pilot.csv')

glimpse(raw_data)
```

# Data Wrangling

```{r}

d <- raw_data %>% 
  # rename column 'education' to 'handedness'
  rename(handedness = education) %>% 
  # deselect columns not needed for analysis 
  select(-picture, -startDate, -startTime, -endTime, -timeSpent, -comments, -gender, -languages)
```

We exclude every individual trial with reaction times below 100 ms and above 1500 ms:

```{r}

d <- d %>% 
  mutate(
    outlier = case_when(RT < 100 ~ TRUE,
                        RT > 1500 ~ TRUE,
                        TRUE ~ FALSE)
  )
```

There are `r d %>% filter(outlier==TRUE) %>% count()` outliers to be excluded.

```{r}

d <- d %>% 
    filter(outlier == FALSE)
```

We exclude all data from practice trials:

```{r}

d <- d %>% 
  filter(trial_type == "main")
```

We only look at correct responses:

```{r}

d <- d %>% 
  filter(correctness == "correct")
```

Given the variables `condition` and `expected`, we create a new variable `correct_response` that will serve as a covariate.

```{r}

d <- d %>% 
mutate(
  correct_response = ifelse(condition == "compatible" & expected == "big", "right", 
                            ifelse(condition == "incompatible" & expected == "small", "right", "left"))
)
```

```{r}

head(d)
```

# Model Specifications

We specify 'compatible' as default level for factor `condition`, and 'right' as default level for factor `correct_response`.

```{r eval=FALSE}

d <- d %>% 
    mutate(
      # Treat predictor variables as unordered factors
      condition = factor(condition, levels = c("compatible", "incompatible")),
      correct_response = factor(correct_response, levels = c("right", "left"))
  )

head(d)
```

## Model 1 (FE)

Model 1 fits a Bayesian linear model with variate `RT` (reaction time) and covariates `condition` (compatible, incompatible) and `correct_response` (right, left). This model does not take an interaction into account.

```{r warnings = FALSE, messages = FALSE, results = 'hide'}

m1_noInteraction <- brm(
  formula = RT ~ condition + correct_response,
  data = d,
  seed = 123
)
```

```{r}

summary(m1_noInteraction)
```

## Model 2 (FE)

Model 2 fits a Bayesian linear model with variate `RT` (reaction time) and covariates `condition` (compatible, incompatible) and `correct_response` (right, left). This model takes an interaction between `condition` and `correct_response` into account.

```{r warnings = FALSE, messages = FALSE, results = 'hide'}

m2_Interaction <- brm(
  formula = RT ~ condition * correct_response,
  data = d,
  seed = 123
)
```

```{r}

summary(m2_Interaction)
```

## Model 3 (RE)

Model 3 fits a Bayesian hierarchical linear model with variate `RT` (reaction time) and covariates `condition` (compatible, incompatible) and `correct_response` (right, left). This model takes an interaction, as well as by-participant random intercepts into account.

```{r warnings = FALSE, messages = FALSE, results = 'hide'}

m3_Interaction_Random_Intercepts <- brm(
  formula = RT ~ condition * correct_response + (1||submission_id),
  data = d,
  iter = 4000,
  control = list(adapt_delta = 0.95),
  seed = 123
)
```

```{r}

summary(m3_Interaction_Random_Intercepts)
```

## Model 4 (RE)

Model 4 fits a Bayesian hierarchical linear model with variate `RT` (reaction time) and covariates `condition` (compatible, incompatible) and `correct_response` (right, left). This model takes an interaction, as well as by-participant random intercepts and slopes into account.

```{r warnings = FALSE, messages = FALSE, results = 'hide'}

m4_Interaction_Random_Intercepts_Slopes <- brm(
  formula = RT ~ condition * correct_response + (condition*correct_response||submission_id),
  data = d,
  iter = 4000,
  control = list(adapt_delta = 0.95),
  seed = 123
)
```

```{r}

summary(m4_Interaction_Random_Intercepts_Slopes)
```

# Model Comparison

In this section, we will compare models 1-4 against each other. Concretely, we will choose the model for analysis based on the "best" posterior predicitve distribution according to the LOO-CV method. We will rely on the `loo_compare` function of the `brms` package and calculate $p$-values using a $z$-score method.

## Model 1 vs. Model 2

```{r}

FE_noInteraction <- loo(m1_noInteraction)
FE_Interaction <- loo(m2_Interaction)

loo_comp_FE <- loo_compare(list(simpler = FE_noInteraction, more_complex = FE_Interaction))
loo_comp_FE
```

```{r}

1 - pnorm(-loo_comp_FE[2,1], loo_comp_FE[2,2])
```

## Model 2 vs. Model 3

```{r}

RE_Intercepts <- loo(m3_Interaction_Random_Intercepts)

loo_comp_FE_vs_RE <- loo_compare(list(simpler = FE_Interaction, more_complex = RE_Intercepts))
loo_comp_FE_vs_RE
```

```{r}

1 - pnorm(-loo_comp_FE_vs_RE[2,1], loo_comp_FE_vs_RE[2,2])
```

## Model 3 vs. Model 4

```{r}

RE_Intercepts_Slopes <- loo(m4_Interaction_Random_Intercepts_Slopes)

loo_comp_RE <- loo_compare(list(simpler = RE_Intercepts, more_complex = RE_Intercepts_Slopes))
loo_comp_RE
```

```{r}

1 - pnorm(-loo_comp_RE[2,1], loo_comp_RE[2,2])
```

## Compare all models

```{r}

loo_compare(list(FE_noInteraction = FE_noInteraction, 
                 FE_Interaction = FE_Interaction, 
                 RE_Intercepts = RE_Intercepts, 
                 RE_Intercepts_Slopes = RE_Intercepts_Slopes))
```
