---
title: "S-R Compatibility Task - Confirmatory Analysis"
author: 
    name: Milan Ewert, Tim Bax, Niklas Bub, Özge Özenoglu
    email: mewert@uos.de, tbax@uos.de, nbub@uos.de, ooezenoglu@uos.de
date: "July 28, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    highlight: tango
---

# Preliminaries

## Load libraries

```{r echo = TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)
library(bayesplot)
# devtools::install_github('michael-franke/bayes_mixed_regression_tutorial/faintr',build_vignettes = TRUE)
library(faintr)

options(mc.cores = parallel::detectCores())

color_palette = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = color_palette)
}

scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = color_palette)
} 
```

## Load raw data 

```{r}
raw_data <- read.csv('https://raw.githubusercontent.com/ooezenoglu/XPlab-2020-SR-Compatibility/master/data/02_main/01-raw-data-main.csv')

glimpse(raw_data)
```

# Optional participant information

```{r}
# check summary statistics regarding the age
raw_data %>% 
  pull(age) %>% 
  summary()
```

```{r message=FALSE, warning=FALSE}
# check counts within the 'gender' variable
raw_data %>%
  group_by(gender) %>% 
  summarise(count = n()/150) %>% 
  mutate(gender = c("no answer", "female", "male"))
```

```{r}
# check comments of participants
raw_data %>% 
  pull(comments) %>% 
  unique()
```

# Data Wrangling

```{r}
d <- raw_data %>% 
  # rename column name 'education' to 'handedness'
  rename(handedness = education) %>% 
  # discard columns not needed for analysis 
  select(-picture, -startDate, -startTime, -endTime, -timeSpent, -comments, -gender, -languages)
```

We exclude every individual trial with reaction times below 100 ms and above 1500 ms:

```{r}
d <- d %>% 
  mutate(
    outlier = case_when(RT < 100 ~ TRUE,
                        RT > 1500 ~ TRUE,
                        TRUE ~ FALSE)
  )
```

There are `r d %>% filter(outlier==TRUE) %>% count()` outliers to be excluded.

```{r}
d <- d %>% 
    filter(outlier == FALSE)
```

We exclude all data from practice trials:

```{r}
d <- d %>% 
  filter(trial_type == "main")
```

We only look at correct responses:

```{r}
d <- d %>% 
  filter(correctness == "correct")
```

Given the variables `condition` and `expected`, we create a new variable `correct_response` that will serve as a covariate.

```{r}
d <- d %>% 
mutate(
  # if the condition is 'compatible' and the square is 'big', store the value 'right' in correct_response
  correct_response = ifelse(condition == "compatible" & expected == "big", "right", 
                            # if the condition is 'incompatible' and the square is 'small', store the value 'right' in correct_response;
                            # else: store the value 'left' in correct_response (case: compatible, small; incompatible, big)
                            ifelse(condition == "incompatible" & expected == "small", "right", "left"))
)
```

```{r}
# peek at the first six rows
head(d)
```

# Summary statistics

```{r}
# show summary statistics about the reaction time
d %>% pull(RT) %>% summary()
```

```{r message=FALSE}
d_agg <- d %>% 
  # look at each level of 'condition' (compatible, incompatible) and 'correct_response' (left, right)
  group_by(condition, correct_response) %>% 
  # get the mean and standard deviation of the reaction time
  summarize(mean_RT = mean(RT),
            sd_RT = sd(RT))

d_agg
```

```{r message=FALSE}
d_agg_participant <- d %>% 
  # look at each level of 'condition' (compatible, incompatible) and 'correct_response' (left, right) for every participant
  group_by(condition, correct_response, submission_id) %>% 
  # get the mean and standard deviation of the reaction time
  summarize(mean_RT = mean(RT),
            sd_RT = sd(RT))

# peek at the first six rows
head(d_agg_participant)
```

```{r fig.align='center'}
# plot the mean reaction times as a function of 'condition' and 'correct_response'
ggplot(d_agg, aes(x = correct_response, y = mean_RT, color = condition)) + 
  geom_point(size = 2) +
  # connect mean reaction times, grouped by 'condition'
  geom_line(aes(group = condition)) +
  xlab("\ncorrect response location") +
  ylab("reaction time in ms\n") +
  theme_classic()
```

```{r fig.align='center'}
# plot the mean reaction times as a function of 'condition' and 'correct_response'
ggplot(d_agg, aes(x = correct_response, y = mean_RT, color = condition)) + 
  geom_point(aes(fill = condition), size = 4, position = position_dodge(width = 0.5)) +
  # add by-participant mean reaction times for each combination of 'condition' and 'correct_response' as semi-transparent points
  geom_point(data = d_agg_participant, aes(x = correct_response, y = mean_RT), size = 2, alpha = 0.2, position = position_dodge(width = 0.5)) +
  xlab("\ncorrect response location") +
  ylab("reaction time in ms\n") +
  theme_classic()
```

# Random effects model

We specify 'compatible' as reference level for factor `condition`, and 'right' as reference level for factor `correct_response`.

```{r}
d <- d %>% 
    mutate(
      # Treat predictor variables as unordered factors
      condition = factor(condition, levels = c("compatible", "incompatible")),
      correct_response = factor(correct_response, levels = c("right", "left"))
  )

# peek at the first six rows
head(d)
```

## Model specification

To answer all of our research hypotheses, we fit one Bayesian hierarchical linear model with variate `RT` (reaction time), covariates `condition` and `correct_response`, and their two-way interaction. Furthermore, we specify by-participant random intercepts. We use the default (flat) priors of the `brms` package for the effect coefficients. 

```{r warnings = FALSE, messages = FALSE, results = 'hide'}
model_RE <- brm(
  formula = RT ~ condition * correct_response + # fixed effects
    (condition * correct_response||submission_id), # random effects
  data = d,
  iter = 3000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  seed = 525
)
```

## Posterior estimates

```{r}
# print model output
model_RE
```

```{r}
# take a look at the chosen priors by brms
prior_summary(model_RE)
```

```{r}
# extract posterior samples
parameter_values <- posterior_samples(model_RE) %>%
  round(3) %>% 
  # select coefficients of interest (population-level effects)
  select(b_Intercept, 
         b_conditionincompatible,
         b_correct_responseleft,
         `b_conditionincompatible:correct_responseleft`) %>% 
  # cast output into tidy format
  pivot_longer(cols = everything(), names_to = "parameter", values_to = "value") %>% 
  select(parameter, value)

# peek at the first six rows
head(parameter_values)
```

```{r}
# store model output in a tibble
parameter_summary <- model_RE %>%
  spread_draws(b_Intercept, 
               b_conditionincompatible,
               b_correct_responseleft,
               `b_conditionincompatible:correct_responseleft`) %>%
  mean_qi()

# extract the mean posterior estimates in a tidy format
mean_values <- parameter_summary %>%  
  pivot_longer(cols = c(b_Intercept, b_conditionincompatible, b_correct_responseleft, 
                        `b_conditionincompatible:correct_responseleft`), 
               names_to = "parameter", 
               values_to = "mean") %>% 
  select(parameter, mean)

# extract the lower bounds of the 95% credible intervals in a tidy format
lowerHDI <- parameter_summary %>% 
  pivot_longer(cols = contains("lower"), 
               names_to = "parameter_lower", 
               values_to = "lower") %>% 
  select(parameter_lower, lower)

# extract the upper bounds of the 95% credible intervals in a tidy format
upperHDI <- parameter_summary %>% 
  pivot_longer(cols = contains("upper"), 
               names_to = "parameter_upper", 
               values_to = "upper") %>% 
  select(parameter_upper, upper)

# combine mean posterior estimates, lower 95% CrI, upper 95% CrI in one tibble
coefficients <- cbind(mean_values, lowerHDI, upperHDI) %>% 
  select(-contains(c("_lower", "_upper")))


coefficients
```

The following plot is inspired by a tutorial of Michael Franke and Timo Röttger (2019)^[Franke, M., & Roettger, T. B. (2019, July 13). Bayesian regression modeling (for factorial designs): A tutorial. https://doi.org/10.31234/osf.io/cdxv3].

```{r fig.align='center'}
# specify new label names for facet_wrap
new_labels <- c("b_Intercept" = "Intercept", 
                "b_conditionincompatible" = "condition: incompatible", 
                "b_correct_responseleft" = "correct response: left", 
                "b_conditionincompatible:correct_responseleft" = "condition: incompatible x correct response: left")

# create a density plot for all parameter values of the coefficients estimated by the model
ggplot(data = parameter_values, mapping = aes(x = value)) +
  facet_wrap(.~parameter, scales = "free", labeller = labeller(parameter = new_labels))  +
  geom_density(fill = "lightgrey") +
  # add mean posterior estimates as dashed, red lines
  geom_vline(data = coefficients, mapping = aes(xintercept = mean), color = "firebrick", lty = "dashed") +
  # mark the 95% credible interall for all coefficients 
  geom_segment(data = coefficients, mapping = aes(x = lower,xend = upper,y = 0,yend = 0), size = 2, color = "firebrick") +
  xlab("\nparameter values") +
  ylab("density\n") +
  theme_classic() +
  theme(strip.background = element_blank())
```

```{r message=FALSE}
# Extract the posterior estimates
derived_posteriors <- model_RE %>%
  spread_draws(b_Intercept, 
               b_conditionincompatible,
               b_correct_responseleft,
               `b_conditionincompatible:correct_responseleft`) %>%
  # Calculate posteriors for each individual level
  mutate(Compatible_right = b_Intercept,
         Compatible_left = b_Intercept + b_correct_responseleft,
         Incompatible_right = b_Intercept + b_conditionincompatible,
         Incompatible_left = b_Intercept + b_conditionincompatible + b_correct_responseleft + 
                              `b_conditionincompatible:correct_responseleft`) %>% 
  # select relevant columns
  select(Compatible_right, Compatible_left, Incompatible_right, Incompatible_left) %>% 
  gather(key = "parameter", value = "posterior") %>%
  # create a new column 'correct_response' (values: right, left)
  # if the current value in 'parameter' contains the string 'right', store 'right' in correct_response; else: 'left'
  mutate(correct_response = ifelse(parameter == "Compatible_right" | parameter == "Incompatible_right",
                        "right", "left"),
         # create a new column 'condition' (values: compatible, incompatible)
         # if the current value in 'parameter' contains the string 'Compatible, store 'compatible' in condition; else: 'incompatible'
          condition = ifelse(parameter == "Compatible_right" | parameter == "Compatible_left",
                        "compatible", "incompatible")) 

# summarize posterior means and 95% Credible Intervals for each combination of condition and correct_response
posteriors_agg <- derived_posteriors %>% 
  group_by(correct_response, condition) %>% 
  summarise(`|95%` = HDInterval::hdi(posterior, credMass = 0.95)[1],
            mean_posterior = mean(posterior),
            `95%|` = HDInterval::hdi(posterior, credMass = 0.95)[2])

posteriors_agg
```

```{r}
posteriors_agg %>% 
  ggplot(aes(x = correct_response, y = mean_posterior, color = condition)) +
  facet_wrap(.~condition) +
  # plot the mean posterior estimates as a function of 'correct_response' and 'condition'
  geom_point(size = 2) +
  # ad the 95% credible intervals for each combination of 'correct_response' and 'condition'
  geom_errorbar(aes(ymin = `|95%`, ymax = `95%|`), width = 0.1) +
  # connect mean posterior estimates, grouped by 'condition'
  geom_line(aes(group = condition)) +
  theme_classic() +
  theme(strip.background = element_blank(), strip.text.x = element_blank()) +
  xlab("\ncorrect response location") +
  ylab("mean posterior\n")
```

# Hypothesis Testing

## Hypothesis 1

> Response times are faster in the compatible mapping condition than in the incompatible mapping condition.

```{r}
compare_groups(
  model = model_RE,
  higher = list("condition" = "incompatible"),
  lower = list("condition" = "compatible")
)
```

## Hypothesis 2

> The compatible S-R mapping yields faster reaction times for right-responses than for left-responses.

```{r}
compare_groups(
  model = model_RE,
  higher = list(condition = "compatible", correct_response = "left"),
  lower = list(condition = "compatible", correct_response = "right")
)
```

## Hypothesis 3

> The incompatible S-R mapping yields slower reaction times for right-responses than for left-responses.

```{r}
compare_groups(
  model = model_RE,
  higher = list(condition = "incompatible", correct_response = "right"),
  lower = list(condition = "incompatible", correct_response = "left")
)
```
