---
title: "S-R Compatibility Task - Exploratory Analysis"
author: 
    name: Milan Ewert, Tim Bax, Niklas Bub, Özge Özenoglu
    email: mewert@uos.de, tbax@uos.de, nbub@uos.de, ooezenoglu@uos.de
date: "July 23, 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    highlight: tango
---

# Preliminaries

## Load libraries

```{r echo = TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
library(brms)
library(tidybayes)
library(bayesplot)
# devtools::install_github('michael-franke/bayes_mixed_regression_tutorial/faintr',build_vignettes = TRUE)
library(faintr)

options(mc.cores = parallel::detectCores())

color_palette = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = color_palette)
}

scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = color_palette)
} 

```

## Load raw data 

```{r}
raw_data <- read.csv('.csv')

glimpse(raw_data)
```

# Data Wrangling

```{r}
d <- raw_data %>% 
  # rename column 'education' to 'handedness'
  rename(handedness = education) %>% 
  # discard columns not needed for analysis 
  select(-picture, -startDate, -startTime, -endTime, -timeSpent, -comments, -gender, -languages)
```

We exclude every individual trial with reaction times below 100 ms and above 1500 ms:

```{r}
d <- d %>% 
  mutate(
    outlier = case_when(RT < 100 ~ TRUE,
                        RT > 1500 ~ TRUE,
                        TRUE ~ FALSE)
  )
```

There are `r d %>% filter(outlier==TRUE) %>% count()` outliers to be excluded.

```{r}
d <- d %>% 
    filter(outlier == FALSE)
```

We exclude all data from practice trials:

```{r}
d <- d %>% 
  filter(trial_type == "main")
```

We only look at correct responses:

```{r}
d <- d %>% 
  filter(correctness == "correct")
```

Given the variables `condition` and `expected`, we create a new variable `correct_response` that will serve as a covariate.

```{r}
d <- d %>% 
mutate(
  # if the condition is 'compatible' and the square is 'big', store the value 'right' in correct_response
  correct_response = ifelse(condition == "compatible" & expected == "big", "right", 
                            # if the condition is 'incompatible' and the square is 'small', store the value 'right' in correct_response;
                            # else: store the value 'left' in correct_response (case: compatible, small; incompatible, big)
                            ifelse(condition == "incompatible" & expected == "small", "right", "left"))
)
```

```{r}
# peek at the first six rows
head(d)
```

# Exploratory Analysis

```{r message=FALSE}
d_agg_expl <- d %>% 
  # look at each level of 'condition', 'correct_response' and 'handedness'
  group_by(condition, correct_response, handedness) %>% 
  # get the mean and standard deviation of the reaction time
  summarize(mean_RT = mean(RT),
            sd_RT = sd(RT))

d_agg_expl
```

```{r message=FALSE}
d_agg_participant_expl <- d %>% 
  # look at each level of 'condition', 'correct_response' and 'handedness' for each participant
  group_by(condition, correct_response, submission_id, handedness) %>% 
  # get the mean and standard deviation of the reaction time
  summarize(mean_RT = mean(RT),
            sd_RT = sd(RT))

head(d_agg_participant_expl)
```

```{r message=FALSE}
handedness_table <- d %>% 
  select(submission_id, handedness) %>% 
  group_by(submission_id, handedness) %>% 
  summarize()

handedness_counts <- tibble(
  handedness = c("right", "left", "I am able to use both hands equally well"), 
  counts = c(sum(handedness_table$handedness=="right"), 
             sum(handedness_table$handedness=="left"), 
             sum(handedness_table$handedness=="I am able to use both hands equally well")))

handedness_counts
```

```{r fig.align='center'}
ggplot(d_agg_expl, aes(x = correct_response, y = mean_RT, color = condition)) + 
  geom_point(size = 2) +
  geom_line(aes(group = condition)) +
  facet_wrap(.~handedness, labeller = labeller(handedness = c("left" = "left-handed", "right" = "right-handed", "I am able to use both hands equally well" = "both-handed"))) +
  xlab("\ncorrect response location") +
  ylab("reaction time in ms\n") +
  theme_classic() +
  theme(strip.background = element_blank(), strip.text = element_text(size = 11))
```

## Model Specification

In case that there are left-handed participants, we fit a Bayesian hierarchical linear model with variate `RT` (reaction time), covariates `condition`, `correct_response`, `handedness`, as well as their three-way interaction. Furthermore, we specify by-participant random intercepts. We use the default (flat) priors of the `brms` package for the effect coefficients. 

```{r warnings = FALSE, messages = FALSE, results = 'hide'}
model_RE <- brm(
  formula = RT ~ condition * correct_response * handedness + # fixed effects
    (condition * correct_response||submission_id), # random effects
  data = d,
  iter = 4000,
  control = list(adapt_delta = 0.99, max_treedepth = 15),
  seed = 525
)
```

##  Posterior estimates

```{r, eval=FALSE}
# print model output
model_RE_expl
```

```{r eval=FALSE}
# take a look at the chosen priors by brms
prior_summary(model_RE_expl)
```

## Hypothesis Testing

> Do right-handers respond faster in the compatible mapping condition than left-handers?

```{r eval=FALSE}
compare_groups(
  model = model_RE_expl,
  higher = list(condition = "compatible", handedness = "left"),
  lower = list(condition = "compatible", handedness = "right")
)
```

> Do left-handers respond faster in the incompatible mapping condition than right-handers?

```{r eval=FALSE}

compare_groups(
  model = model_RE_expl,
  higher = list(condition = "incompatible", handedness = "right"),
  lower = list(condition = "incompatible", handedness = "left")
)
```

> Do left-handers respond faster to left-responses than to right-responses?

```{r eval=FALSE}
compare_groups(
  model = model_RE_expl,
  higher = list(correct_response = "right", handedness = "left"),
  lower = list(correct_response = "left", handedness = "left")
)
```

> Do right-handers respond faster to right-responses than to left-responses?

```{r eval=FALSE}
compare_groups(
  model = model_RE_expl,
  higher = list(correct_response = "left", handedness = "right"),
  lower = list(correct_response = "right", handedness = "right")
)
```
